---
title: "Descriptive Statistics"
author: "Aleksandr Fisher"
output: beamer_presentation
theme: metropolis
latex_engine: xelatex
highlight: zenburn
---


## Plan

1. Overview of Last Week
2. Measurement 
3. Descriptive Statistics
4. Readings
5. Wrap up

## Measurement

* Social science is about developing and testing causal theories:
  + Does minimum wage change levels of employment?
  + Does outgroup contact influence views on immigration?
* Theories are made up of concepts:
  + Minimum wage, level of employment, outgroup contact, views on
immigration.
  + We took these for granted when talking about causality.
* Important to consider how we measure these concepts.
  + Some more straightforward: what is your age?
  + Others more complicated: what does it mean to “be liberal”?
  + Have to create an operational definition of a concept to make it into a
variable in our dataset.

## Example

- Concept: presidential approval.
- Conceptual definition:
  - Extent to which US adults support the actions and policies of the current US
president.
  - Operational definition:
- “On a scale from 1 to 5, where 1 is least supportive and 5 is more supportive,
how much would you say you support the job that Donald Trump is doing as
president?”

## Measurement Error

- **Measurement error**: chance variation in our measurements.
  - individual measurement = exact value + chance error
  - chance errors tend to cancel out when we take averages.
- No matter how careful we are, a measurement could have always come out
differently.
  - Panel study of 19,000 respondents: 20 reported being a citizen in 2010 and
then a non-citizen in 2012.
  - Data entry errors.
- **Bias**: systematic errors for all units in the same direction.
  - individual measurement = exact value + bias + chance error.
  - “Did you vote?” ~ overreporting

## Definitions

- A **variable** is a series of measurements about some concept.
- **Descriptive statistics** are numerical summaries of those measurements.
  - If we smart enough, we wouldn’t need them: just look at the list of numbers
and completely understand.
- Two salient features of a variable that we want to know:
  - **Central tendency**: where is the middle/typical/average value.
  - **Spread** around the center: are all the data close to the center or spread out?

## Center of the Data

- "Center" of the data: Typical/average value
- **Mean**: sum of the values divided by the number od observations
- **Median**: the "middle" of a sorted list of numbers.
- Median more robust to outliers
  - Example 1: data = {0, 1, 2, 3, 5}, mean = 2.2, median = 2
  - Example 2: data = {0, 1, 2, 3, 100}, mean = 21.2, median = 2
  
## Spread of the data

- Are the data close to the center?
- **Range**: [min(x), max(x]
- **Quantile** (quartile, quintile, percentile, etc):
  - 25th percentile = lower quartile (25% of the data below this value)
  - 50th percentile = median (50% of the data below this value)
  - 75th percentile = upper quartile (75% of the data below this value)
- **Interquartile range (IQR)**: a measure of variability
  - How spread out is the middle half of the data?
  - Is most of the data really close to the median or are the values spread out?
- One definition of outliers: over 1.5 × IQR above the upper quartile or below
lower quartile.

## Standard deviation

- **Standard deviation**: On average, how far away are data points from the
mean?
$$\sqrt\frac{\sum((x - \bar{x})^{2})}{n-1}$$
- Steps:

  1. Subtract each data point by the mean
  2. Square each resulting difference
  3. Take the sum of these values
  4. Divide by n-1
  5. Take the square root

## Variance

- **Variance** = standard deviation (squared)

$$\frac{\sum((x - \bar{x})^{2})}{n-1}$$

## How large is large?

- Need a way to put any variable on common units.
-  **z-score**: $$\frac{x - \bar{x}}{\sigma}$$

- Interpretation:
  - Positive values above the mean, negative values below the mean
  - Units now on the scale of standard deviations away from the mean
  - Intuition: data more than 3 SDs away from mean are rare.
  
## Data

```{r cars, messages=FALSE, warning=FALSE, background='#F7F7F7'}
summary(cars)
```
