---
name: Aleksandr Fisher
# position: "Data Analyst, College Board"
# aboutme: Ph.D. Political Science
address: ""
phone: 267 337 5263
email: "aleksandr.jr.fisher@gmail.com"
twitter: aleksandrfisher
github: AleksandrFish
linkedin: aleksandrfisher
headcolor: 414141
website: aleksandrfisher.com
date: "`r format(Sys.time(), '%B %Y')`"
output: 
  vitae::awesomecv
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(vitae)
library(tidyr)
library(dplyr)
library(readr)
library(glue)
library(tibble)
library(lubridate)
library(here)
library(scales)
# jobs <- read_csv(here::here("data", "jobs.csv")) %>%
#   fill(Start, End, What, With, Where, Tag) %>%  
#   mutate(Start = mdy(Start),
#          End = mdy(End)) %>% 
#   mutate(When = case_when(
#     is.na(End) ~ glue("{year(Start)}--Present") %>% as.character(),
#     year(Start) == year(End) ~ year(End) %>% as.character(),
#     TRUE ~ glue("{year(Start)}--{year(End)}") %>% as.character()
#   ))
# 
# funding <- read_delim(here::here("data", "funding.txt"), delim = "\t")
# 
# education <- read_delim(here::here("data", "education.txt"), delim = "\t") %>%
#   fill(Start, End, What, With, Where, Tag) %>%  
#   mutate(Start = mdy(Start),
#          End = mdy(End)) %>% 
#   mutate(When = case_when(
#     is.na(End) ~ glue("{year(Start)}--Present") %>% as.character(),
#     year(Start) == year(End) ~ year(End) %>% as.character(),
#     TRUE ~ glue("{year(Start)}--{year(End)}") %>% as.character()
#   ))
# 
# pubs <- read_delim(here::here("data", "pubs.txt"), delim = "\t")
```



# Education
```{r edu}
# education %>% 
#   arrange(desc(Start)) %>% 
#   detailed_entries(
#     what = What,
#     when = When,
#     with = With,
#     where = Where,
#     why = Why
#   )
```

**<span style="font-size:1em;">George Washington University</span>** \hfill _Washington, DC_  
Ph.D. in Political Science  \hfill _2019_

**<span style="font-size:1em;">George Washington University</span>** \hfill _Washington, DC_  
M.A in Political Science  \hfill _2017_

**<span style="font-size:1em;">Temple University</span>** \hfill _Philadelphia, PA_  
B.A. in History and Political Science \hfill _2014_


# Academic Positions
```{r teaching}
# jobs %>% 
#   arrange(desc(Start)) %>% 
#   detailed_entries(
#     what = What,
#     when = When,
#     with = With,
#     where = Where,
#     why = Why
#   )
``` 

**<span style="font-size:1em;">Haverford College</span>** \hfill _Haverford, PA_  
Assistant Professor \hfill _Sept. 2020 - Dec. 2020_  

- Designed and instructed _Introduction to Quanitative Methods for Political Science_ which included
lessons on causal inference, regression, and data manipulaton/visualization in **R** (Undergraduate students)

**<span style="font-size:1em;">Jefferson University</span>** \hfill _Philadelphia, PA_  
Instructor \hfill _Jan. 2019 - May 2019_  

- Designed and instructed two sections of _Debating Global Issues_, which included lessons on economic globalization, development, and regime change (Undergraduate students)
 

**<span style="font-size:1em;">George Washington University</span>** \hfill _Washington, DC_  
Instructor \hfill _Sept. 2018 - August. 2018_  

- Designed and instructed a summer course _Science, Technology, and Politics_, which explored the influence of AI on political processes
(Undergraduate students)
- Led a 4-day course reviewing fundamental math principles including probability and hypothesis tests; instruction in basic **R** (Doctoral students)
- Teaching assistant for _Introduction to Empirical Political Analysis_ (Doctoral students)
- Teaching assistant for _Nonlinear Models_ (Doctoral students)
- Led discussion sections for _Introduction to Comparative Politics_ (Undergraduate students)

# Non-Academic Experience
```{r jobs}
# jobs %>% 
#   arrange(desc(Start)) %>% 
#   detailed_entries(
#     what = What,
#     when = When,
#     with = With,
#     where = Where,
#     why = Why
#   )
```

**<span style="font-size:1em;"> The College Board </span>** \hfill _Newtown, PA_  
Data Analyst \hfill _Jan. 2020 - Present_  

 - Developed SAS and Python code that increased productivity and accuracy of the quality management processes related to the SAT, PSAT 10, and PSAT 8/9
 - Built a web application in Python/Flask to automate the creation of excel spreadsheets that tracks how College Board products align with state standards, reduced process time by dozens of hours every month
 - Created new archives or archive features that added significant value to psychometric work
 

**<span style="font-size:1em;">Insight Data Science</span>** \hfill _New York, NY_  
Data Science Fellow \hfill _Sept. 2019 - Nov 2019_  

- Built an end-to-end product in Python/Flask for an early-stage startup in three weeks that predicts
respondents’ severity of depression from text inputs using a random forest model
- Cleaned unstructured text data using Pandas and NLTK, created predictive features using tf-idf and sentiment
analysis, and compared machine learning models using scikit-learn
- Predicted people’s level of depression within 2 points on a 16-point clinically verified scale, doubling the
accuracy of a baseline model
 

# Peer Reviewed Publications
```{r pubs}
# pubs %>% 
#   arrange(desc(date)) %>% 
#   detailed_entries(
#     what = title,
#     when = date,
#     with = authors,
#     where = journal,
#     why = paste(volume, pages)
#   ) 
# tmp <- vector(mode = "list", length = nrow(pubs))
# 
# for (i in seq(1, length(tmp))) {
#   a1 <- pubs[i, ]
#   if (a1$pages == " ") {
#     tmp[[i]] <- paste0(a1$authors, 
#                   " (",
#                   a1$date,
#                   "), ", 
#                   a1$title, 
#                   ", ", 
#                   a1$journal, 
#                   ", ", 
#                   a1$volume)
#     
#   } else {
#     tmp[[i]] <- paste0(a1$authors, 
#                   " (",
#                   a1$date,
#                   "), ", 
#                   a1$title, 
#                   ", ", 
#                   a1$journal, 
#                   ", ", 
#                   a1$volume, 
#                   ", ",
#                   a1$pages)
#   }
# }
# 
# tmp <- do.call(args = tmp, what = rbind)
# 
# pubs %>% 
#   tibble::add_column(., tmp = tmp) %>%
#   arrange(desc(date)) %>% 
#   detailed_entries(
#     what = tmp
#   )
# pubs %>% 
#   dplyr::mutate(., tmp = paste0(authors, 
#                   " (",
#                   date,
#                   "), ", 
#                   title, 
#                   ", ", 
#                   journal, 
#                   ", ", 
#                   volume, 
#                   ", ", 
#                   pages)) %>%
#   arrange(desc(date)) %>% 
#   detailed_entries(
#     what = tmp
#   )
```

- Fisher, Aleksandr. "Demonizing the enemy: The influence of Russian state-sponsored media on American audiences." Post-Soviet Affairs (2020): 1-16.[[Link](https://www.tandfonline.com/doi/abs/10.1080/1060586X.2020.1730121)]

- Fisher, Aleksandr. "Trickle Down Soft Power: Do Russia's Ties to European Parties Influence Public Opinion?." _Foreign Policy Analysis_ (2020). [[Link](https://academic.oup.com/fpa/advance-article-abstract/doi/10.1093/fpa/oraa013/5825379)]

- Fisher, Aleksandr. "A New Cold War? International Public Opinion of Russia and the United States." _International Journal of Public Opinion Research_ 32.1 (2020): 143-152. [[Link](https://academic.oup.com/ijpor/article-abstract/32/1/143/5397055)]

- Evers, Miles M., Aleksandr Fisher, and Steven D. Schaaf. "Is There a Trump Effect? An Experiment on Political Polarization and Audience Costs." _Perspectives on Politics_ 17.2 (2019): 433-452. [[Link](https://www.cambridge.org/core/journals/perspectives-on-politics/article/is-there-a-trump-effect-an-experiment-on-political-polarization-and-audience-costs/BDD30AB0D76FD935B43952121A49C15A)]

```{r policypubs}
# pubs %>% 
#   arrange(desc(date)) %>% 
#   detailed_entries(
#     what = title,
#     when = date,
#     with = authors,
#     where = journal,
#     why = paste(volume, pages)
#   ) 
# tmp <- vector(mode = "list", length = nrow(pubs))
# 
# for (i in seq(1, length(tmp))) {
#   a1 <- pubs[i, ]
#   if (a1$pages == " ") {
#     tmp[[i]] <- paste0(a1$authors, 
#                   " (",
#                   a1$date,
#                   "), ", 
#                   a1$title, 
#                   ", ", 
#                   a1$journal, 
#                   ", ", 
#                   a1$volume)
#     
#   } else {
#     tmp[[i]] <- paste0(a1$authors, 
#                   " (",
#                   a1$date,
#                   "), ", 
#                   a1$title, 
#                   ", ", 
#                   a1$journal, 
#                   ", ", 
#                   a1$volume, 
#                   ", ",
#                   a1$pages)
#   }
# }
# 
# tmp <- do.call(args = tmp, what = rbind)
# 
# pubs %>% 
#   tibble::add_column(., tmp = tmp) %>%
#   arrange(desc(date)) %>% 
#   detailed_entries(
#     what = tmp
#   )
# pubs %>% 
#   dplyr::mutate(., tmp = paste0(authors, 
#                   " (",
#                   date,
#                   "), ", 
#                   title, 
#                   ", ", 
#                   journal, 
#                   ", ", 
#                   volume, 
#                   ", ", 
#                   pages)) %>%
#   arrange(desc(date)) %>% 
#   detailed_entries(
#     what = tmp
#   )
```

# Policy Publications

- If Trump backs down from his latest Iran threat, will he lose support? We checked. _Washington Post, Monkey Cage_, 2019, with Miles M. Evers and Steven Schaaf. [[Link](https://www.washingtonpost.com/politics/2019/06/26/if-trump-backs-down-his-latest-iran-threat-will-he-lose-support-we-checked/)]

- Perceptions of Russian Interference in U.S. Elections Matters as Much as Actual Involvement. _Foreign Policy Research Institute_, 2019. [[Link](https://www.fpri.org/article/2019/05/perceptions-of-russian-interference-in-u-s-elections-matter-as-much-as-the-actual-involvement/)]

- Moldova and Ukraine: Diverging Neighbors? _Foreign Policy Research Institute_, 2017. [[Link](https://www.fpri.org/article/2017/02/moldova-ukraine-diverging-neighbors/)]

- Ukraine’s Struggle for Democracy: Severe Challenges Ahead. _Foreign Policy Research Institute_, 2015, with Amb. Adrian Basora. [[Link](https://www.fpri.org/article/2014/07/ukraines-struggle-for-democracy-severe-challenges-ahead/)]

- Putin’s Greater Novorossiya: The Dismemberment of Ukraine. _Foreign Policy Research Institute_, 2015, with Amb.Adrian Basora. [[Link](https://www.fpri.org/docs/basora-_putins_greater_novorossiya.pdf)]

- Home-Grown Terrorism: Russia’s Ticking Bomb. _ISN, Center for Security Studies (CSS)_, 2015, with Maia Otarashvili. [[Link](https://www.files.ethz.ch/isn/187916/ISN_181348_en.pdf)]

- Ukraine Crisis Timeline as of January 30, 2015. _Foreign Policy Research Institute_, 2015. [[Link](https://www.fpri.org/2015/02/ukraine-crisis-timeline-as-of-january-30-2015/)]

# Invited Workshops


- American Political Science Association, _Washington, DC_ \hfill _2019_  

- Midwest Political Science Association Meeting, _Chicago, IL_ \hfill _2017-2019_  

- 5th Student Conference for Trajectories of Change, _Berlin, Germany_ \hfill _2019_

- Annual DC Area Comparative Workshop, _George Washington University_ \hfill _2019_

- Junior Scholar’s Symposium, Cato Institute \hfill _2018_

- Comparative Politics Workshop, _George Washington University_ \hfill _2018_

- Annual DC Area Comparative Workshop, _Georgetown University_ \hfill _2018_

- ASN Convention, _Columbia University_ \hfill _2018_

- Regime Evolution, Institutional Change, and Social Transformation in Russia, _Yale University_ \hfill _2018_

- Post-Communist Workshop, _George Washington University_ \hfill _2017_

- Rethinking Russian Media Strategy and Influence, _George Washington University_ \hfill _2017_

- Post-Communist Workshop, _George Washington University_ \hfill _2016_

# Awards

- Summer Dissertation Fellowship, _George Washington University_ \hfill _2019_ 

- Dean’s Dissertation Completion Fellowship, _George Washington University_ \hfill _2019_ 

- Trajectories of Change’ Fieldwork Grant, _ZEIT-Stiftung Ebelin und Gerd Bucerius_ \hfill _2019_ 

- Bradley Fellowship, _The Lynde and Harry Bradley Foundation_ \hfill _2018_

- Hoffman Dissertation Fellowship, _Institute for European, Russian, and Eurasian Studies_ \hfill _2018_

- Shirley H. and Robert L. Richards Endowment Scholarship, _George Washington University_ \hfill _2017_

- Rumsfeld Foundation Graduate Fellow, _Rumsfeld Foundation_ \hfill _2017_

- Humane Studies Fellowship, _Institute for Humane Studies_ \hfill _2017_

- Graduate Research Fund, _George Washington University_  \hfill _2016_

- Michael Hoffman Fund, _Institute for European, Russian, and Eurasian Studies_ \hfill _2016_

- Professor Edward R. and Essie Baron Memorial, _Temple University_ \hfill _2014_

- Ronald C. Schwarzkopf Endowed Award in Political Science, _Temple University_ \hfill _2014_



# Technical Skills

**Languages**: Python, R, Stata, SAS, SQL, Latex,

**Tools**: Pandas, NumPy, Scikit-Learn, NLTK, Spacy, Beautifulsoup, Matplotlib, Seaborn, Flask 

**Quantitative Methods**: Survey experiments, natural experiments, machine learning, decision trees, random forest, natural language processing, cluster analysis, various time-series cross-sectional models, causal mediation analysis, multinomial logistic regression, factor analysis, web scraping 
