---
title: "Analysis of Burt's IQ Data"
author: "John Fox"
date: "`r as.character(Sys.Date())`"
output: html_document
---


```{r echo=FALSE}
# include this code chunk as-is to set options
knitr::opts_chunk$set(comment=NA, prompt=TRUE, fig.height=8, fig.width=8) 
#    use fig.height=6.5, fig.width=6.5 for word output
```

## A Graph of Burt's Data


```{r}
library(car)
Burt$class <- with(Burt, factor(class, levels=c("low", "medium", "high")))
scatterplot(IQbio ~ IQfoster | class,  data=Burt, 
            boxplots=FALSE, smooth=FALSE)
```

The regression lines for the three classes are not very different, but they are
also remarkably parallel.

## Burt's Regression


```{r}
burt.mod <- lm(IQbio ~ IQfoster*class, data=Burt)
summary(burt.mod)
Anova(burt.mod)
```

There is no evidence of an interaction, and only weak evidence of a class main effect.
There is very strong evidence for an IQfoster "effect."

## Why Should We Strongly Suspect that the Data Are Cooked?

The p-value for the interaction is very close to 1, indicating that the three
regression lines are *almost perfectly* parallel. Data this parallel would occur by
chance less than 1 percent of the time, even if there were no interaction
in the population.
