---
title: "Fibonacci Numbers"
author: "John Fox"
date: "`r as.character(Sys.Date())`"
output: html_document
---


```{r echo=FALSE}
# include this code chunk as-is to set options
knitr::opts_chunk$set(comment=NA, prompt=TRUE, fig.height=8, fig.width=8) 
#    use fig.height=6.5, fig.width=6.5 for word output
```

## Calculating Fibonacci Numbers by Recursion

```{r}
fib0 <- function(n, verbose=FALSE){
    if (verbose) cat("n = ", n, " ")
    if (n <= 2) return(1)
    fib0(n - 1, verbose=verbose) + fib0(n - 2, verbose=verbose)
}
sapply(1:10, fib0)

```

I've included a verbose argument to trace the recursive function calls (see below).

## Calculating Fibonacci Numbers by Iteration

```{r}
fib1 <- function(n){
    if (n <= 2) return(1)
    last.minus.1 <- 1
    last.minus.2 <- 1
    for (i in 3:n){
        save <- last.minus.1
        last.minus.1 <- last.minus.1 + last.minus.2
        last.minus.2 <- save
    }
    last.minus.1
}

sapply(1:10, fib1)
```

## Computing Fibonacci Numbers by Binet's Formula

```{r}
fib2 <- function(n){
    round(((1 + sqrt(5))/2)^n/sqrt(5))
}
sapply(1:10, fib2)

```

## Comparing the Efficiency of the three Solutions

```{r}
options(scipen=10)  # suppress scientific notation
system.time(print(fib0(30)))
system.time(print(fib1(70)))
system.time(print(fib2(70)))
```

The latter two solutions are too fast to time with `system.time()` but we can get timings with the **microbenchmark** package:

```{r}
library(microbenchmark) # need to install the microbenchmark package
microbenchmark(fib1=fib1(70),
               fib2=fib2(70))
```

As expected `fib2()`, which uses Binet's formula is much faster (proportionally) than `fib1()`, which uses iteration.

`fib2(70)` produces the right answer, but not `fib2(71)`:

```{r}
fib1(71)
fib2(71)
```

To see why the recursive function `fib0()` is so inefficient, trace the computation:

```{r}
fib0(8, verbose=TRUE)
```

It's apparent that some of the same Fibonacci numbers (e.g., for n = 2) end up being computed over and over and over again.
