---
title: "Experimental Design in R"
output: beamer_presentation
---

```{r setup, include=FALSE}
opts_chunk$set(cache=TRUE)
```

# Intro to Experimental Design

* Planning
    + Dependent Variable = outcome
    + Independent Variable(s) = explanatory

* Design
* Analysis

# Load Data
Remove all previous data with:
rm(list = ls())

```{r , echo = TRUE}
# Load the ToothGrowth dataset
data(ToothGrowth)
```

# T.Test
```{r}
# Perform a two-sided t-test
t.test(x = ToothGrowth$len, 
       alternative = "two.sided", mu = 18)
```

# T.Test by Treatment

* Recall that randomization of subjects in an experiment helps spread any natural variability between subjects evenly across groups.

* In the experiment that yielded the ToothGrowth dataset, guinea pigs were randomized to receive Vitamin C either through orange juice or ascorbic acid, indicated in the dataset by the supp variable. It's natural to wonder if there is a difference in tooth length by supplement type - a question that a t-test can also answer!

```{r}
# Perform a t-test
ToothGrowth_ttest <- t.test(len ~ supp, 
                            data = ToothGrowth)

# Load broom
library(broom)
```

# T.Test by Treatment (2)


```{r}
# Tidy the t-test model object
tidy(ToothGrowth_ttest)
```