---
title: "Introduction to Data in R"
output: pdf_document
---


##Introduction
```{r}
library(openintro)
data(hsb2)
summary(hsb2)
str(hsb2)
```

```{r}
library(dplyr)
glimpse(hsb2)
```

##Types of Variables
Categorical data are often stored as factors in R. 

Recall from the video that the filter() function from dplyr allows you to filter a dataset to create a subset containing only certain levels of a variable. 

# Subset of emails with big numbers: email50_big
```{r}
hsb_big <- hsb2 %>%
  filter(gender=="male")
```

# Glimpse the subset
```{r}
glimpse(hsb_big)
```
# Table of gender variable
```{r}
table(hsb_big$gender)
```
# Drop levels
```{r}
hsb_big$gender <- droplevels(hsb_big$gender)
```
# Another table of number variable
```{r}
table(hsb_big$gender)
```
#Load Email data
```{r}
data(email50)
```
# Calculate median number of characters: med_num_char
```{r}
med_num_char <- median(email50$num_char)
```
# Create num_char_cat variable in email50
```{r}
email50 <- email50 %>%
  mutate(num_char_cat = ifelse(num_char < med_num_char, "below median", "at or above median"))
```  
# Count emails in each category
```{r}
table(email50$num_char_cat)
```
# Create number_yn column in email50
```{r}
email50 <- email50 %>%
  mutate(number_yn= ifelse(number=="none", "no", "yes"))
```
# Visualize number_yn
```{r}
ggplot(email50, aes(x = number_yn)) +
  geom_bar()
```
# Load packages
```{r}
library(dplyr)
library(tidyr)
```
# Count number of male and female applicants admitted
```{r}
hsb_race <- hsb2 %>%
  count(gender, race)
```
# View result
```{r}
print(hsb_race)
```
# Spread the output across columns
```{r}
hsb_race %>%
  spread(gender, n)
```  

# Table of counts of admission status and gender
 count(schtyp, gender) %>%
# Spread output across columns based on admission status
 spread(schtyp, n)  %>%
# Create new variable
mutate(Perc_type = public/ (public+private))
print(hsb_type)
# Table of counts of admission status and gender for each department
hsb_type2 <- hsb2 %>%
  count(ses, schtyp, gender) %>%
  spread(schtyp, n) %>%
# Percentage of those admitted to each department
  mutate(Perc_type = public / (public+private))
  print(hsb_type2)

library(openintro)
data(county)

county_noDC<- county %>%
  filter(state !="District of Columbia") %>%
  droplevels()

#Simple Random Sample  
county_srs <- county_noDC %>%
  sample_n(size=150)
  glimpse(county_srs)
  
county_srs %>%
  group_by(state) %>%
  count()
  
#Stratified Sample
county_str <- county_noDC %>%
  group_by(state) %>%
  sample_n(size=3)
  glimpse(county_str)
  
#Beauty in the Classroom
download.file("http://www.openintro.org/stat/data/evals.RData", destfile = "evals.RData")
load("evals.RData")

# Inspect variable types
glimpse(evals)
str(evals) # Another option

# Remove non-factor variables from this vector
cat_vars <- c("rank", "ethnicity", "gender", "language",
              "cls_level", "cls_profs", "cls_credits",
              "pic_outfit", "pic_color") 
              
# Recode cls_students as cls_type: evals
evals <- evals %>%
  # Create new variable
  mutate(cls_type = factor(ifelse(cls_students <= 18, "small",
                      ifelse(cls_students >= 19 & cls_students <= 59, "midsize",  "large"))))
                      
# Scatterplot of score vs. bty_avg
ggplot(evals, aes(x=bty_avg, y=score)) +
  geom_point()
  
# Scatterplot of score vs. bty_avg colored by cls_type
ggplot(evals, aes(x=bty_avg, y=score, color=cls_type)) +
  geom_point()