---
title: "States Regression Exercises"
author: "John Fox"
date: "`r as.character(Sys.Date())`"
output: html_document
---

```{r echo=FALSE}
# include this code chunk as-is to set options
knitr::opts_chunk$set(comment=NA, prompt=TRUE, fig.height=8, fig.width=8) 
#    use fig.height=6.5, fig.width=6.5 for word output
```

## Regression of SAT Math Scores on Teachers' Pay


```{r}
library(car)
mod.1 <- lm(SATM ~ pay, data=States)
summary(mod.1)
```
As teacher's pay goes up, on average SAT math scores go down!

## Regression of SAT Math Scores on Both Teachers' Pay and Percent Taking the Exam

```{r}
mod.2 <- lm(SATM ~ pay + percent, data=States)
summary(mod.2)
compareCoefs(mod.1, mod.2)
```

Holding percent taking the exam constant, there's a *positive* relationship of SAT math scores to teachers' pay. On the other hand, holding teachers' pay constant, SAT math scores are negatively related to percent of students taking the exam.

## Graphing the Data

We probably should have started here! 

Here's a scatterplot matrix for the three variables:

```{r}
scatterplotMatrix(~ SATM + pay + percent, data=States, span=0.9)
```

Both SATM and even more so percent taking the exam are bimodal. Some of the marginal relationships are apparently nonlinear, suggesting that there may be problems in the linear regression.

Some Diagnostics
----------------

```{r, fig.height=4}
crPlots(mod.2, span=0.9)
avPlots(mod.2)
```
```{r}
qqPlot(mod.2)
```

Indeed, nonlinearity seems an issue, though not an extreme issue. On the other hand, the added-variable plot doesn't suggest the presence of influential cases, nor is the distribution of the studentized residuals particularly prolematic.
